---
title       : Analysis of Transmission Factors on MPG
subtitle    : 
author      : Trey Beeman
job         : Data Scientist
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Executive Summary 

We will examine the `mtcars` data with a goal to fit a linear model to determine if automatic or manual transmission is better for fuel efficiency ("mpg") and we will quantify the difference should it exist.

--- .class #id 

## Data Exploration and Preparation

```{r preparation} 
data(mtcars); mtcars$am <- factor(mtcars$am)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$cyl <- factor(mtcars$cyl); mtcars$carb <- factor(mtcars$carb)
```
Data Variables: mileage (`mpg`), number of cylinders (`cyl`), 
displacement (`disp`), horse power (`hp`), rear axle ratio (`drat`), 
weight (`wt`), 1/4 mile time (`qsec`), cylinder arrangement (`vs`) , transmission type (`am`), number of forward gears (`gear`) and number of carburetors (`carb`) for 32 different vehicle models.

---

## Str() of the Data

```{r str of mtcars}
str(mtcars)
```

---

## Boxplot Automatic v. Manual Transmission

```{r data exploration, echo = FALSE}
library(ggplot2)
g <-  ggplot(mtcars, aes(x = am, y = mpg)) + geom_boxplot(aes(fill = am)) +
        labs(title = "Figure 1: MPG in Automatic and Manual Transmission Vehicles", 
             x = "Transmission Type") + 
        scale_x_discrete(labels = c("Automatic", "Manual")) +
        scale_fill_discrete("Type", labels = c("Automatic","Manual"))
g
```

---

## Model 1: Automatic Transmission v. Manual and MPG

```{r mpg analyis}
mpg_fit1 <- lm(mpg ~ am, mtcars)
summary(mpg_fit1)$r.squared
```

$R^{2}$ indicates this model does deficient in explaining the variance in the data around its mean.  

---

## Model 2: Step function to find most important variables

We will use the `step()` function to find a model with an improved $R^{2}$ value.  The `step()` function will indicate the best fit using the AIC method.  It will check each variable in the model to find one with the best fit according to the AIC value.  The final model is below.

```{r best, results='hide'}
best <- step(lm(mpg~.,mtcars), direction = "both")
```

```{r best fit model}
summary(best)$call
mpg_fit2 <- lm(mpg ~ wt + qsec + am, mtcars)
```

---

## Results from the Step() Function

```{r summary best, echo = FALSE}
summary(mpg_fit2)
```

---

## Comments on the results of the Step() function

```{r adj R Squared best fit model}
summary(mpg_fit2)$adj.r.squared
```

The model returned by the `step()` function (above), gives an Adjusted $R^{2}$ indicating that this model explains 83% of the variance of the data around its mean, so this model is much stronger than the earlier fit.  Also, by comparing the residual plots in Figure 2 and Figure 3, we can see there is more heteroskedacity in this model.  Figure 3 indicates several outliers  Chrysler Imperial, Fiat 128 and Toyota Corolla which we may wish to exclude.

---

## ANOVA on the models

Coefficients are statistically significant compared to the model that does not include them.

```{r anova}
mpg_fit1.a <- lm(mpg ~ am + qsec, mtcars)
anova(mpg_fit1, mpg_fit1.a, mpg_fit2)
```

---

## Residual Plot 1

```{r residual 1, echo = FALSE}
plot(mpg_fit1, which = 1, main = "Figure 2: Residual Plot fit1")
```

---

## Residual Plot 2

```{r residual 2, echo = FALSE}
plot(mpg_fit2, which = 1, main = "Figure 3: Residual Plot fit2")
```

---

## Comments on Residual Plots

1. Comparing Figure 2 and Figure 3 - more heteroskedacity in the best fit from the `Step()` function.

2. Figure 3 indicates several outliers Chrysler Imperial, Fiat 128 and Toyota Corolla which we may wish to exclude.

---

### Conclusions

```{r coeff summaries}
summary(mpg_fit2)$coefficients
```

1.)Manual transmission contributes to lower gas mileage

2. The `am1` factor is significant at $\alpha = .05$ confirming the relationship.

3. With weight (`wt`) and 1/4 mile time (`qsec`) held constant, average of 2.9358 increase in mpg when using a manual transmission.

4. Significant with 95% confidence.

---
